The Host Galaxies of Rapidly Evolving Transients in the Dark Energy Survey

### 2020-05-14 RET_Hosts_CWR.tex
Updates from Overleaf

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
%Co-authors:
\newcommand{\phil}[1]{\color{red}#1 \color{black}}
\newcommand{\mike}[1]{\color{cyan}#1 \color{black}}
\newcommand{\replychris}[1]{\color{magenta}#1 \color{black}}
\newcommand{\replylluis}[1]{\color{brown}#1 \color{black}}
\newcommand{\comment}[1]{\textbf{[#1]}} %comments

@@

% Abstract of the paper
\begin{abstract}
Rapidly evolving transients (RETs), also termed fast blue optical transients (FBOTs), are a distinct class of astrophysical event. They are characterised by lightcurves that decline much faster than standard supernovae (SNe), span vast ranges in peak luminosity and can be seen to redshifts greater than 1. Their evolution on fast timescales has hindered high quality follow-up observations, such that their origin and explosion/emission mechanism remains unexplained. In this paper we define the largest sample of RETs to date, comprising 106 objects from the Dark Energy Survey, and perform the first substantial analysis of RET host galaxies. Using deep-stacked photometry and emission-lines from OzDES spectroscopy, we derive stellar masses and star-formation rates (SFRs) for 49 host galaxies, and metallicities for 37. We find that RETs explode exclusively in star-forming galaxies and are thus likely associated with massive stars. Comparing RET hosts to samples of host galaxies of other explosive transients as well as field galaxies, we find that RETs prefer galaxies with high specific SFRs, indicating a link to young stellar populations, similar to stripped-envelope SNe. RET hosts appear to show a lack of chemical enrichment, their metallicities akin to long duration gamma-ray bursts and superluminous SN host galaxies. There are no clear relationships between properties of the host galaxies and the peak magnitudes or decline rates of the transients themselves.

@@

In the standard paradigm of stellar evolution, stars with a zero-age main sequence (ZAMS) mass above $8M_{\sun}$ are believed to explode as a result of a catastrophic collapse of their iron cores and are known as core-collapse supernovae (CCSNe). CCSNe can be split into observationally-determinded subclasses based on their lightcurve and spectral evolution: SNe II display hydrogen features in their spectra \citep{Minkowski1941}, and are thought to occur in stars that retain a large fraction of their hydrogen envelope. Conversely, SNe Ib and Ic do not show signatures of hydrogen \citep[e.g.][]{Filippenko1997} and are thus referred to collectively as stripped-envelope SNe (SESNe), having undergone a partial removal of their outer atmospheres. The SN IIb subclass, which shows hydrogen only at early epochs that disappears after a few weeks \citep{Filippenko1988}, is also commonly grouped along with SESNe. SNe IIn, with the n standing for \textit{narrow}, which refers to the apparent width of the hydrogen emission lines when compared to standard SNe II \citep{Schlegel1990}, indicating their origin from the ejecta impacting on slow-moving circumstellar material (CSM). Since the turn of the century, observations of CCSNe, whose lightcurves are primarily powered by the radioactive decay of freshly synthesised Ni-56, have been supplemented by rarer, more exotic transient classes.

 Long duration gamma-ray bursts (LGRBs), although first discovered in the 1960s \citep{Klebesadel1973}, were only unequivocally linked to collapsing massive stars through their associations with broad-lined type Ic SNe \citep{Galama1998,Hjorth2003}. Thought to be caused by accretion onto a newly-formed black hole at the centre of a collapsing, rapidly-rotating massive star \citep[e.g.][]{Woosley1993,Woosley2006a,Woosley2006b}, LGRBs comprise roughly $1\%$ of all SNe Ic, which themselves make up only 15\% of all CCSNe \citep{Kelly2012,Graham2016}. The second exotic class of SNe is the particularly bright superluminous supernovae (SLSNe; e.g. \citealt{Quimby2011, Gal-Yam2012}). Originally grouped due to their slowly-evolving lightcurves and extreme luminosity (peaking at $M_B < -21$~mag; 10-100 times brighter than regular CCSNe), recent observations have revealed a continuum of spectroscopically similar objects with peaks as faint as $M_B \sim -19$~mag \citep{DeCia2018,Lunnan2018,Angus2019}, similar to the bright end of the CCSN luminosity function \citep{Li2011}. The lightcurve evolution of SLSNe is not well described by models of Ni-56 decay, with the most popular alternative hypothesis being the magnetic coupling of the ejecta with the spin down of a newly formed, rapidly rotating magnetar.

 Along with observations of the transients themselves, host galaxies are frequently-used laboratories from which strong inferences about the progenitor stars and explosion mechanisms can be made. CCSNe are confined almost exclusively to galaxies hosting recent or ongoing star formation, due to their origin from massive stars. There are correlations between the expected progenitor mass of different sub-classes of CCSNe and host galaxy properties. On average, SESNe reside in galaxies with higher specific star-formation rates (sSFRs; \citealt{James2006,Kelly2008}), while studies of the local environments tend to show that SESNe explode closer to \HII~regions than SNe II, indicating that the progenitors are younger and more massive than the various sub-classes of hydrogen-rich SNe II  \citep[e.g.][]{Anderson2012,Galbany2018}. More extreme events tend to occur in galaxies low in mass and high in sSFR, with both GRBs \citep[e.g.][]{Fruchter2006,LeFloch2006,Levesque2010,Kruehler2015,Vergani2015,Perley2016b,Palmerio2019,Taggart2019} and to an even greater degree SLSNe \citep[e.g.][]{Neill2011,Lunnan2014,Leloudas2015,Angus2016,Schulze2018,Taggart2019} exhibiting this association.

 The chemical composition of the interstellar medium (ISM) is an important consideration when comparing host galaxy properties. While it does not appear to play a significant role in the relative production of CCSNe (although there are some trends, with SESNe typically found in slightly less metal-rich environments than SNe II; \citealt{Galbany2018}), it appears to be vitally important in the production of LGRBs and SLSNe. Theory predicts that the production of a LGRB should only be possible in stars with a metallicity of $Z/Z_{\sun}\leq 0.3$ \citep{Woosley1993}  in order for the likely Wolf-Rayet or blue supergiant progenitors not to lose their outer atmospheres through metal-driven winds, thus conserving sufficient angular momentum to power the black-hole-driven jet or rapidly rotating magnetar. Many LGRB host galaxy studies have indeed revealed a metallicity threshold to be observed between 0.5 and 1 times the solar value \citep[e.g.][]{Stanek2006,Modjaz2008,Kruehler2015,Perley2016b,Japelj2016,Vergani2017}, although this is not a rigid cut.  
SLSN host galaxies also appear to be lower in metallicity than would be expected for their stellar mass, with a suppression of SLSN production at a value around half-solar \citep{Lunnan2014,Chen2016a,Perley2016c}. They also require a particularly high sSFR, suggesting that they are explosions of very young, rapidly rotating massive stars.

Recently, inspection of high-cadence, large-area survey data sets have revealed yet more exotic transients that are less easy to explain with conventional models. \citet{Drout2014} presented a sample of rapidly evolving transients (RETs; also termed `Fast Blue Optical Transients' - FBOTs or `Fast Evolving Luminous Transients' - FELTs) in the Pan-STARRS survey (PS1). \citet{Arcavi2016} reported three rapidly rising, highly luminous objects in the Supernova Legacy Survey (SNLS) and one in the Palomar Transient Factory (PTF), of which one (SNLS04D4ec) declines rapidly like the PS1 sample. \citet[hereafter P18]{Pursiainen2018} expanded the known number of RETs to beyond 80 with their sample from the Dark Energy Survey (DES), spanning a redshift range of $\sim 0$ to $>1$. A further sample of five objects has been discovered by the Hyper Suprime-Cam Subaru Strategic Program (SSP) Transient Survey \citep{Tampo2020}. RETs typically rise to peak brightness in less than 10 days, and decline to 10\% of their peak brightness within 30 days, much faster than typical SNe. The photometric measurements of the PS1 and DES RETs seem to be well described by simple expanding blackbodies, although a handful show declining photospheric radii from the outset. Due to the rapid nature of their lightcurves and location at high-redshift, spectral coverage is sparse and signal-to-noise ratio (SNR) is low, such that there has not yet been a conclusive detection of absorption or emission features from the transients and thus the physical mechanism responsible for their rapid evolution remains unexplained.

There are a limited number of events detected in the local Universe whose properties are consistent with the RETs seen in the samples of PS1 and DES at cosmological distances, the most widely studied of which is AT2018cow \citep[e.g.][]{Prentice2018,Margutti2019,Perley2019}. The transient declined from its discovery, with constraints on the rise time of 1 day, and across the full range of observed wavelengths did not resemble any known SN, GRB afterglow, or kilonova (KN; \citealt{Ho2019}). There are many diverse explanations for the power source of AT2018cow touted in the literature, including: magnetars \citep{Mohan2020}; electron capture collapse of merged white dwarfs \citep{Lyutikov2019}; a tidal disruption event (TDE) of a white dwarf \citep{Kuin2019} or of a main sequence star by an intermediate mass black hole \citep{Perley2019}; common envelope jets supernova (CEJSN; \citealt{Soker2019}); or a wind-driven transient \citep{Uno2020}.
Other nearby rapid transients include the local fast-declining SN-like transient SN2018kzr \citep{McBrien2019} which is explained by the destruction of a white dwarf, and KSN-2015K \citep{Rest2018} whose fast rise and decline is explained by the shock of an SN running into previously-expelled material.  It is currently unclear whether these transients do indeed represent the local analogues of the DES and PS RETs.

In this paper, we present the first comprehensive study of the host galaxies of RETs. We make use of the final DES sample, which builds on \citetalias{Pursiainen2018} by adding the 5th and final year of DES-SN observations as well as more refined discovery techniques. Using the deep DES photometry from \citet[hereafter W20]{Wiseman2020} and spectra from OzDES \citep{Lidman2020} we derive host galaxy properties in order to compare them to samples of CCSNe, LGRBs, and SLSNe, as well as the individual local rapid transients. For clarity, we will use the term RET to refer only to events in the high-redshift samples of DES and PS1.

@@

The original method of finding RETs in the DES-SN data (and presented in \citetalias{Pursiainen2018}) was designed to be simple and used light curve modelling with Gaussian and linear fits. The simplistic method made it possible to look for exotic transients without knowing their observed characteristics beforehand and resulted in a large sample of photometrically selected fast transients. However, as the search was simplistic and relied heavily on visual inspection of the available data (etc. images, light curves, host galaxy information), it is impossible to be sure how complete the sample is. For instance, due to the large redshift range within the sample it is entirely possible that distant events could have been missed due to time dilation. Here, a more sophisticated search method is presented. As only a fraction of transients in DES-SN have redshift information from their host galaxies, the search must be done in the observer frame.  The key features of RETs that separate them from most traditional SNe types are the fast light curve evolution and blue colour at peak. Even though both of these quantities depend on the redshift of the transient, they still seem to distinguish the the fast events from traditional SN types.

To assess the unique photometric properties of RETs, we take the sample of 72 from the \citetalias{Pursiainen2018} method, updated with 20 extra objects found using that method in the fifth season of DES, along with spectroscopically confirmed SNe of types Ia, Ibc, II, and SLSNe from DES that pass the following criteria:
\begin{enumerate}
\item The transient was only detected in one DES-SN observing season.
\item Maximum observed brightness in both $g$- and $r$-bands was brighter than 24 mag (in the eight `shallow' DES-SN fields) or 25 mag (in the two `deep' fields).
\item $g$- and $r$-band observations used for the colour had to be taken within 2 days of each other in the observer frame.
\end{enumerate}

Of objects passing these cuts, the RETs cluster at shorter timescales and bluer colours than other SNe, even in the observer frame. However, using photometric data points directly has several problems that can be improved. For one, measuring peak colour is problematic: DES-SN did not always observe $g$- and $r$-bands on the same or even consecutive nights, thus making it impossible to measure the peak colour in a number of cases. Measuring rise times of 10-15 days is difficult to perform with a one week cadence when it has to be done without fitting a light curve model, hence the rise time values are spread over a wide range. To negate this issue, we use lightcurves interpolated using Gaussian Processes (GP) as presented in \citet{Pursiainen2020}. The interpolated light curves have a 0.5 day cadence and every epoch and band has a flux value, and an associated uncertainty. Using this technique, SNe Ia and RETs populate two distinct regions of $g-r$ vs. $t_{\mathrm{rise; r}}$ parameter space (Fig. \ref{fig:selection}), where $t_{\mathrm{rise; r}}$ is the time to rise from non-detection to peak $r$-band brightness, and the $g-r$ colour is measured at peak brightness. We define a region in this parameter space which minimises the contamination of non-RETs (purity) while maximising the total fraction of RETs (completeness). The resulting limits are $-1.5 < g-r < 0.6 $ and $t_{\mathrm{rise; r}} <16.5$, and the parameter space can be seen in Fig. \ref{fig:selection}.

We process all $\sim 30,000$ DES-SN transient candidates with GP. In order to reduce the contamination from active galactic nuclei (AGN), we use a basic convolutional neural network (CNN) classifier. We train the CNN on spectroscopically confirmed SNe of all types and on spectroscopically typed AGN, and use it to separate the sample into two photometric subtypes: AGN-like and SNe-like. The classifier returns SNe-like objects with an accuracy of 0.992 on the test set; the remaining AGN are removed by manual vetting.
The SNe passing the CNN classifier are subjected to LC quality cuts, resulting in 2259 objects, of which 939 lie inside the colour and rise-time region which we defined as RETs. These objects are subject to a further set of cuts. We impose a cut based on a fit of the LC with the \texttt{PSNID} software \citep{Sako2008}. We use thresholds of \texttt{FITPROB}$<0.91$ and \texttt{PBAYES}$<0.82$ to remove highly-probably SNe Ia, which removes 46 objects from the RET parameter space. In order to further remove longer-lived SNe, the decline time to half of the peak brightness must be $<24$ days. This removes 347 SNe, resulting in 546 objects remaining inside the parameter space. The final 546 transients have been visually inspected, with the majority rejected for being spurious detections, obvious multi-season variability that was not picked up by the CNN, or showing evidence for a longer timescale decline.

Using the above method recovers 87 of the 92 RETs found using the \citetalias{Pursiainen2018} technique, and adds a further 14. The five were not recovered as their GP lightcurves were fainter than the limits given above in either g or r band.
We refer to the resulting sample as DES RETs. Of the 106 objects in the sample, 96 have a host galaxy detected in deep host galaxy photometry of \citetalias{Wiseman2020} when using the Directional Light Radius method \citep{Sullivan2006} to associate hosts as per \citetalias{Wiseman2020}. Of these, 49 have a host galaxy spectroscopic redshift which we access through an internal release of the OzDES Global Redshift Catalog (GRC; v.2020\_01\_04). The full OzDES redshift catalogue will be available alongside the public data release detailed in \citet{Lidman2020}. A further three have redshifts obtained from narrow lines observed in spectra of the transients themselves. We do not consider these three objects for the analysis, since we are unable to separate transient and host contributions to the spectra.
A selection of the host galaxies can be seen in Fig. \ref{fig:mosaic}, centred on the location of the transient. The figure showcases the diversity of host galaxy morphologies and colours, while also displaying the limitations of ground-based observations of high-redshift, relatively small galaxies in terms of spatial resolution.
Fig. \ref{fig:z_dist} shows the distribution of redshifts amongst the 49 hosts for which such a measurement was possible. The effect that the redshift selection function has on the results is discussed in Section \ref{subsec:disc_bias}.
The observational properties of the 96 detected hosts are displayed in Table \ref{tab:obs}. We highlight objects that were not presented in \citetalias{Pursiainen2018}. We also highlight a small subset of objects for which the redshift has taken on a new value than that presented in \citetalias{Pursiainen2018} due to further OzDES observations leading to a more accurate determination.

@@

To compare with CCSNe, we draw on the untargetted sample of 47 SNe II from the Palomar Transient Factory (PTF; \citealt{Stoll2013}), which is likely complete in terms of hosts (all SNe have an associated host). While this sample lies at much lower redshift than the DES RETs (a maximum of 0.18 and mean of 0.05), redshift evolution is easier to account for than unknown incompleteness. We add to this the compilation of 56 untargeted SESNe from \citet{Sanders2012}, with a maximum redshift of 0.26 and a mean of 0.05. Since \citet{Sanders2012} do not report host galaxy magnitudes, stellar masses or SFRs, we cross-match the SN positions with the the Sloan Digital Sky Survey (SDSS; \citealt{York2000}) Data Release 16 (DR16; \citealt{Ahumada2019}) and perform our own SED fit using the method outlined in Section \ref{subsec:sedfit}. We are able to do this for 38 objects, with the others lying outside of the SDSS footprint.

@@

\begin{figure}
\includegraphics[width=0.5\textwidth]{figs/RET_BPT.png}
\caption{Baldwin-Phillips-Terlovich (BPT) diagram for RET hosts, showing that the emission lines are consistent with being generated by star formation rather than AGN activity. The three curves show the delimitation between star formation and AGN at $z =0, 0.5, 1$ according to \citet{Kewley2013}.
\label{fig:bpt}}
\end{figure}

>>>>

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
%Co-authors:
\newcommand{\phil}[1]{\color{red}#1 \color{black}}
\newcommand{\mike}[1]{\color{cyan}#1 \color{black}}
\newcommand{\replyrick}[1]{\color{magenta}#1 \color{black}}
\newcommand{\replylluis}[1]{\color{brown}#1 \color{black}}
\newcommand{\comment}[1]{\textbf{[#1]}} %comments

@@

% Abstract of the paper
\begin{abstract}
Rapidly evolving transients (RETs), also termed fast blue optical transients (FBOTs), are a distinct class of astrophysical event. They are characterised by lightcurves that decline much faster than common classes supernovae (SNe), span vast ranges in peak luminosity and can be seen to redshifts greater than 1. Their evolution on fast timescales has hindered high quality follow-up observations, and thus their origin and explosion/emission mechanism remains unexplained. In this paper we define the largest sample of RETs to date, comprising 106 objects from the Dark Energy Survey, and perform the first substantial analysis of RET host galaxies. Using deep-stacked photometry and emission-lines from OzDES spectroscopy, we derive stellar masses and star-formation rates (SFRs) for 49 host galaxies, and metallicities for 37. We find that RETs explode exclusively in star-forming galaxies and are thus likely associated with massive stars. Comparing RET hosts to samples of host galaxies of other explosive transients as well as field galaxies, we find that RETs prefer galaxies with high specific SFRs, indicating a link to young stellar populations, similar to stripped-envelope SNe. RET hosts appear to show a lack of chemical enrichment, their metallicities akin to long duration gamma-ray bursts and superluminous SN host galaxies. There are no clear relationships between properties of the host galaxies and the peak magnitudes or decline rates of the transients themselves.

@@

In the standard paradigm of stellar evolution, stars with a zero-age main sequence (ZAMS) mass above $8M_{\sun}$ are believed to explode as a result of a catastrophic collapse of their iron cores and are known as core-collapse supernovae (CCSNe). CCSNe can be split into observationally-determinded subclasses based on their lightcurve and spectral evolution: SNe II display hydrogen features in their spectra \citep{Minkowski1941}, and are thought to occur in stars that retain a large fraction of their hydrogen envelope. Conversely, SNe Ib and Ic do not show signatures of hydrogen \citep[e.g.][]{Filippenko1997} and are thus referred to collectively as stripped-envelope SNe (SESNe), having undergone a partial removal of their outer atmospheres. The SN IIb subclass, which shows hydrogen only at early epochs that disappears after a few weeks \citep{Filippenko1988}, is also commonly grouped along with SESNe. \replyrick{SNe IIn display much narrower hydrogen emission lines when compared to standard SNe II \citep{Schlegel1990}. The narrow emission originates from the ejecta impacting on slow-moving circumstellar material (CSM)}. Since the turn of the century, observations of CCSNe, whose lightcurves are primarily powered by the radioactive decay of freshly synthesised Ni-56, have been supplemented by rarer, more exotic transient classes.

 Long duration gamma-ray bursts (LGRBs), although first discovered in the 1960s \citep{Klebesadel1973}, were unequivocally linked to collapsing massive stars through their associations with broad-lined type Ic SNe \citep{Galama1998,Hjorth2003}. Thought to be caused by accretion onto a newly-formed black hole at the centre of a collapsing, rapidly-rotating massive star \citep[e.g.][]{Woosley1993,Woosley2006a,Woosley2006b}, LGRBs comprise roughly $1\%$ of all SNe Ic, which themselves make up only 15\% of all CCSNe \citep{Kelly2012,Graham2016}. \replyrick{Another} exotic class of SNe is the particularly bright superluminous supernovae (SLSNe; e.g. \citealt{Quimby2011, Gal-Yam2012}). Originally grouped due to their slowly-evolving lightcurves and extreme luminosity (peaking at $M_B < -21$~mag; 10-100 times brighter than \replyrick{typical} CCSNe), recent observations have revealed a continuum of spectroscopically similar objects with peaks as faint as $M_B \sim -19$~mag \citep{DeCia2018,Lunnan2018,Angus2019}, \replyrick{which overlaps with} the bright end of the CCSN luminosity function \citep{Li2011}. The lightcurve evolution of SLSNe is not well described by models of Ni-56 decay, with the most popular alternative hypothesis being the magnetic coupling of the ejecta with the spin down of a newly formed, rapidly rotating magnetar.

 Along with observations of the transients themselves, \replyrick{studies of} host galaxies are frequently \replyrick{used to make strong inferences about the progenitor stars and explosion mechanisms}. CCSNe are confined almost exclusively to galaxies hosting recent or ongoing star formation, due to their origin from massive stars. There are correlations between the expected progenitor mass of different sub-classes of CCSNe and host galaxy properties. On average, SESNe reside in galaxies with higher specific star-formation rates (sSFRs; \citealt{James2006,Kelly2008}), while studies of the local environments tend to show that SESNe explode closer to \HII~regions than SNe II, indicating that the progenitors are younger and more massive than the various sub-classes of hydrogen-rich SNe II  \citep[e.g.][]{Anderson2012,Galbany2018}. More extreme events tend to occur in galaxies low in mass and high in sSFR, with both GRBs \citep[e.g.][]{Fruchter2006,LeFloch2006,Levesque2010,Kruehler2015,Vergani2015,Perley2016b,Palmerio2019,Taggart2019} and to an even greater degree SLSNe \citep[e.g.][]{Neill2011,Lunnan2014,Leloudas2015,Angus2016,Schulze2018,Taggart2019} exhibiting this association.

 The chemical composition of the interstellar medium (ISM) is an important consideration when comparing host galaxy properties. While it does not appear to play a significant role in the relative production of CCSNe (although there are some trends, with SESNe typically found in slightly less metal-rich environments than SNe II; \citealt{Galbany2018}), it appears to be vitally important in the production of LGRBs and SLSNe. Theory predicts that the production of a LGRB should only be possible in stars with a metallicity of $Z/Z_{\sun}\leq 0.3$ \citep{Woosley1993}  in order for the likely Wolf-Rayet or blue supergiant progenitors not to lose their outer atmospheres through metal-driven winds, thus conserving sufficient angular momentum to power the black-hole-driven jet or rapidly rotating magnetar. Many LGRB host galaxy studies have indeed revealed a metallicity threshold between 0.5 and 1 times the solar value \citep[e.g.][]{Stanek2006,Modjaz2008,Kruehler2015,Perley2016b,Japelj2016,Vergani2017}, although this is not a rigid \replyrick{threshold}.  
SLSN host galaxies also appear to be lower in metallicity than would be expected for their stellar mass, with a suppression of SLSN production \replyrick{above a value of about} half-solar \citep{Lunnan2014,Chen2016a,Perley2016c}. \replyrick{Like LGRBs, SLSNe} also require a particularly high sSFR, suggesting that they are explosions of very young, rapidly rotating massive stars.

Recently, inspection of high-cadence, large-area survey data sets have revealed more exotic transients that are \replyrick{difficult} to explain with conventional models. \citet{Drout2014} presented a sample of rapidly evolving transients (RETs; also termed `Fast Blue Optical Transients' - FBOTs or `Fast Evolving Luminous Transients' - FELTs) in the Pan-STARRS survey (PS1). \citet{Arcavi2016} reported three rapidly rising, highly luminous objects in the Supernova Legacy Survey (SNLS) and one in the Palomar Transient Factory (PTF), of which one (SNLS04D4ec) declines rapidly like the PS1 sample. \citet[hereafter P18]{Pursiainen2018} expanded the known number of RETs to beyond 80 with their sample from the Dark Energy Survey (DES), spanning a redshift range of $\sim 0$ to $>1$. A further sample of five objects has been discovered by the Hyper Suprime-Cam Subaru Strategic Program (SSP) Transient Survey \citep{Tampo2020}. RETs typically rise to peak brightness in less than 10 days, and decline to 10\% of their peak brightness within 30 days, much faster than typical SNe. The photometric measurements of the PS1 and DES RETs seem to be well described by expanding blackbodies, although a handful show declining photospheric radii from the \replyrick{first detection}. Due to the rapid nature of their lightcurves and location at high-redshift, spectral coverage is sparse and signal-to-noise ratio (SNR) is low, such that there has not yet been a conclusive detection of absorption or emission features from the transients and thus the physical mechanism responsible for their rapid evolution remains unexplained.

There are a limited number of events detected in the local Universe whose properties appear consistent with the RETs seen at cosmological distances in the PS1 and DES samples, the most widely studied of which is AT2018cow \citep[e.g.][]{Prentice2018,Margutti2019,Perley2019}. The transient declined from its discovery, with constraints on the rise time of 1 day, and \replyrick{from X-rays through to radio wavelengths} did not resemble any known SN, GRB afterglow, or kilonova (KN; \citealt{Ho2019}) . There are many diverse explanations for the power source of AT2018cow touted in the literature, including: magnetars \citep{Mohan2020}; electron capture collapse of merged white dwarfs \citep{Lyutikov2019}; a tidal disruption event (TDE) of a white dwarf \citep{Kuin2019} or of a main sequence star by an intermediate mass black hole \citep{Perley2019}; common envelope jets supernova (CEJSN; \citealt{Soker2019}); or a wind-driven transient \citep{Uno2020}.
Other nearby rapid transients include the local fast-declining SN-like transient SN2018kzr \citep{McBrien2019} which is explained by the \replyrick{accretion-induced collapse of a white dwarf or a white dwarf-neutron star merger}, and KSN-2015K \citep{Rest2018} whose fast rise and decline is explained by the shock of an SN running into previously-expelled material.  It is currently unclear whether these transients represent the local analogues of the DES and PS\replyrick{1} RETs.

In this paper, we present the first comprehensive study of the host galaxies of RETs. We make use of the final DES sample, which builds on \citetalias{Pursiainen2018} by adding the 5th and final \replyrick{season} of DES-SN observations as well as more refined \replyrick{selection} techniques. Using the deep DES photometry from \citet[hereafter W20]{Wiseman2020} and spectra from OzDES \citep{Lidman2020} we derive host galaxy properties \replyrick{and} compare them to \replyrick{samples of host galaxies of} CCSNe, LGRBs, and SLSNe, as well as the individual local rapid transients. For clarity, we will use the term RET to refer only to events in the high-redshift samples of DES and PS1.

@@

The original method of finding RETs in the DES-SN data (and presented in \citetalias{Pursiainen2018}) was designed to be simple and used light curve modelling with Gaussian and linear fits. The simplistic method made it possible to look for exotic transients without knowing their observed characteristics beforehand and resulted in a large sample of photometrically selected fast transients. However, as the search was simplistic and relied heavily on visual inspection of the available data (etc. images, light curves, host galaxy information), it is impossible to \replyrick{quantify the completeness of the sample}. For instance, due to the large redshift range within the sample it is entirely possible that distant events \replyrick{with longer rise times could have been missed due to time dilation stretching their lightcurves, while faster events and low redshift may evolve on timescales quicker than the survey cadence}. Here, a more sophisticated search method is presented. As only a fraction of transients in DES-SN have redshift information from their host galaxies, \replyrick{we perform the search} in the observer frame. The key features of RETs that separate them from most traditional SNe types are the fast light curve evolution (rise to peak in $\lesssim15$ days) and blue colour at peak ($g-r \lesssim 0.5$). Even though both of these quantities depend on the redshift of the transient, they still effectively distinguish the the fast events from traditional SN types. We thus attempt to select a sample based on observed rise times and colours at peak brightness.

\subsubsection{Gaussian processed lightcurves}
Using observed photometric data points directly to infer rise time and colour has several problems that can be improved. For one, measuring peak colour is problematic: DES-SN did not always observe $g$- and $r$-bands on the same or even consecutive nights in the `deep' fields, thus making it impossible to measure the peak colour accurately. Measuring rise times of 10-15 days is difficult to perform with a one week cadence when it has to be done without fitting a lightcurve model. To negate this issue, we use interpolate the lightcurves of all DES-SN candidates using Gaussian Processes (GP) as presented in \citet{Pursiainen2020}. The interpolated lightcurves have a 0.5 day cadence and every epoch and band has a flux value and an associated uncertainty.

\subsubsection{Photometric definition of RETs \label{subsubsec:trise_gr}}
To make an improved selection of RETs we use a parameter space described by observed $g-r$ colour and the time taken to rise from non-detection to peak $r$-band magnitude in the observer frame. Using rise times and colours from GP lightcurves, we populate this parameter space with the sample of 72 RETs from the \citetalias{Pursiainen2018} method, updated with 20 extra objects found using that method in the fifth season of DES.\replyrick{ We add spectroscopically confirmed SNe of types Ia, Ibc, II, and SLSNe observed by DES in order to verify that they are rejected by the search method}. We keep objects passing the following criteria and disregard those that fail:
\begin{enumerate}
\item The transient was detected in \replyrick{only} one DES-SN observing season.
\item Maximum observed brightness in both $g$- and $r$-bands was brighter than 24 mag (in the eight `shallow' DES-SN fields) or 25 mag (in the two `deep' fields), as in \citetalias{Pursiainen2018}.
%\item $g$- and $r$-band observations used for the colour had to be taken within 2 days of each other in the observer frame.
\end{enumerate}

 SNe Ia and RETs populate two distinct regions of $g-r$ vs. $t_{\mathrm{rise; r}}$ parameter space (Fig. \ref{fig:selection}), where $t_{\mathrm{rise; r}}$ is the time to rise from non-detection to peak $r$-band brightness, and the $g-r$ colour is measured at peak brightness. RETs appear bluer and faster than the typical SNe. We define a region in this parameter space which minimises the contamination of non-RETs (purity) while maximising the total fraction of RETs (completeness). The resulting limits are $-1.5 < g-r < 0.6 $ and $t_{\mathrm{rise; r}} <16.5$, \replyrick{corresponding to the red box in Fig. \ref{fig:selection}}.

\subsubsection{Removal of active galactic nuclei \label{subsubsec:cnn}}

In order to apply selection criterion i) from Section \ref{subsubsec:trise_gr} it is necessary to distinguish between DES-SN candidates that are truly multi-season events (typcially active galactic nuclei; AGN) and those that are single-season events with spurious detections in other seasons. 
To detect and filter out AGN we use a basic convolutional neural network (CNN) classifier. We train the CNN on spectroscopically confirmed SNe of all types as well as the 92 RETs identified using the \citetalias{Pursiainen2018} method (181 objects) and on spectroscopically typed AGN (182 objects), and use it to separate the sample into two photometric subtypes: AGN-like and SNe-like. The classifier returns SNe-like objects with an accuracy of 0.992 on the test set (391 SNe and 79 AGN). No SNe were classified as AGN-like. The remaining AGN classified as SNe-like are removed by manual vetting later in the process.

\subsubsection{Final DES RET sample}
The GP lightcurves of all DES-SN candidates are classified as AGN-like or SNe-like by the CNN, and are then subject to the lightcurve quality cuts i) and ii) (Section \ref{subsubsec:trise_gr}), resulting in 2259 objects, of which 939 lie inside the colour and rise-time region which we defined \replyrick{for} RETs. These objects are subject to a further set of cuts \replyrick{in order to remove remaining contaminants}. We impose a cut based on a fit of the lightcurve with the \texttt{PSNID} software \citep{Sako2008}. \replyrick{\texttt{PSNID} returns a normalised goodness of fit to different SN templates, along with a probability of it being a SN Ia given that it is a SN. To remove highly-probably SNe Ia, we use thresholds of $<0.91$ and $<0.82$ respectively to the above algorithms,} which removes 46 objects from the RET parameter space. In order to further remove longer-lived SNe, the decline time to half of the peak brightness must be $<24$ days. This removes 347 SNe, resulting in 546 objects remaining inside the parameter space. The final 546 transients have been visually inspected by MP, with the majority rejected for being spurious detections, obvious multi-season variability that was not picked up by the CNN, or showing evidence for a longer timescale decline.

Using the above method recovers 87 of the 92 RETs found using the \citetalias{Pursiainen2018} technique, and adds a further 14. The five were not recovered as their GP lightcurves were fainter than the limits given above in either g or r band.
We refer to the resulting sample as DES RETs. Of the 106 objects in the sample, 96 have a host galaxy detected in deep host galaxy photometry of \citetalias{Wiseman2020} when using the Directional Light Radius method \citep{Sullivan2006} to associate hosts as per \citetalias{Wiseman2020}. Of these, 49 have a host galaxy spectroscopic redshift which we access through an internal release of the OzDES Global Redshift Catalog (GRC; v.2020\_01\_04). The full OzDES redshift catalogue will be available alongside the public data release detailed in \citet{Lidman2020}. A further three have redshifts obtained from narrow lines observed in spectra of the transients themselves. We do not consider these three objects for the analysis, since we are unable to separate transient and host contributions to the spectra.
A selection of the host galaxies \replyrick{is shown} in Fig. \ref{fig:mosaic}, centred on the location of the transient. The figure showcases the diversity of host galaxy morphologies and colours, while also displaying the limitations of ground-based observations of high-redshift, relatively small galaxies in terms of spatial resolution.
Fig. \ref{fig:z_dist} shows the distribution of redshifts amongst the 49 hosts for which such a measurement was possible. The effect that the redshift selection function has on the results is discussed in Section \ref{subsec:disc_bias}.
The observational properties of the 96 detected hosts are displayed in Table \ref{tab:obs}. We highlight objects that were not presented in \citetalias{Pursiainen2018}. We also highlight a small subset of objects for which the redshift has taken on a new value than that presented in \citetalias{Pursiainen2018} due to further OzDES observations leading to a more accurate determination.

@@

To compare with CCSNe, we draw on the untargetted sample of 47 SNe II from the Palomar Transient Factory (PTF; \citealt{Stoll2013}), which is likely complete in terms of hosts (all SNe have an associated host). While this sample lies at much lower redshift than the DES RETs (a maximum of 0.18 and mean of 0.05), redshift evolution is easier to account for \replyrick{in a less biased way than correcting for }unknown incompleteness. We add to this the compilation of 56 untargeted SESNe from \citet{Sanders2012}, with a maximum redshift of 0.26 and a mean of 0.05. Since \citet{Sanders2012} do not report host galaxy magnitudes, stellar masses or SFRs, we cross-match the SN positions with the the Sloan Digital Sky Survey (SDSS; \citealt{York2000}) Data Release 16 (DR16; \citealt{Ahumada2019}) and perform our own SED fit using the method outlined in Section \ref{subsec:sedfit}. We are able to do this \replyrick{SED fit} for 38 objects, with the others lying outside of the SDSS footprint.

@@

\begin{figure}
\includegraphics[width=0.5\textwidth]{figs/RET_BPT.png}
\caption{Baldwin-Phillips-Terlovich (BPT) diagram for RET hosts, showing that the emission lines are consistent with being generated by star formation rather than AGN activity. The three curves show the delimitation between star formation and AGN at $z =0, 0.5, 1$ according to \citet{Kewley2013}. \replyrick{The redshift of the RETs is indicated by the colourbar on the right.}
\label{fig:bpt}}
\end{figure}
